---
title: "data_import"
author: "Joe LaRocca"
date: "2024-09-17"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(haven)
```

## Import the FAS Litters CSV

Import the dataset and clean the names.

```{r}

litters_df <- read_csv(file = "data_import_examples/FAS_litters.csv")
litters_df <- janitor::clean_names(litters_df)

```

## Look at the dataset

```{r}

litters_df

head(litters_df)
tail(litters_df, 10)

```


```{r eval = FALSE}

view(litters_df)

```

## Skipping Data

Not necessary here, but if there are "data rules" in the first few lines, use "skip = n" 
within the read_csv command.

## Missing Data

Use the "na" parameter to specify what you want to be listed as "NA" so that R reads number variables
as numbers as opposed to characters.

```{r}

litters_df <- 
  read_csv(
    file = "data_import_examples/FAS_litters.csv",
    na = c("NA", "", ".")                     
  )

litters_df <- janitor::clean_names(litters_df)

```
## Calculate Mean of a Column of a Tibble

```{r}

mean(pull(litters_df, gd0_weight), na.rm = TRUE)

```

## Coding "Group" as a Factor Variable

```{r}

litters_df <- 
  read_csv(
    file = "data_import_examples/FAS_litters.csv",
    na = c("NA", "", "."),
    col_types = cols(
      Group = col_factor()
    )
  )

litters_df <- janitor::clean_names(litters_df)

```


## Import the FAS Pups CSV

```{r}

pups_df <- read_csv(file = "data_import_examples/FAS_pups.csv")

```

## Import an Excel File: MLB 2011 Summary Data (readxl package necessary)

```{r}

mlb_df <- read_excel(
  "data_import_examples/mlb11.xlsx", 
  sheet = "mlb11"
  ) 

```

## Import a SAS file: Public Pulse Data (haven package necessary)

```{r}

pulse_df <- read_sas("data_import_examples/public_pulse_data.sas7bdat")

```

## Never use read.csv(), use read_csv() only

```{r}

litters_df_bad <- read.csv("data_import_examples/FAS_litters.csv")
head(litters_df_bad, 10)

# tibble command read_csv also requires you to make exact references to columns
# don't use dollar signs -- not necessary to take column out of data frame

```

